### 2013-03-22 doc/sosp13/tasks.tex
touchups
commit to merge

The basic unit of execution in Grappa is a {\em task}. Each task is
represented by a function pointer and its arguments. A new task is enqueued at
spawn time; when resources are free, it is allocated a stack, bound to a core,
and executed.

During execution, a task yields control of the processor core whenever it
performs a long-latency operation, allowing the processor to remain busy while
waiting for the operation to complete. In addition, a programmer can direct
scheduling explicitly via the Grappa API calls shown in
Figure~\ref{fig:scheduling}. To minimize yield overhead, the Grappa scheduler
operates entirely in user-space and does little more than store register state
of one task and load that of another. Context switch times are as low as 38ns
even when switching amongst thousands of tasks.
\TODO{We should expand on the cleverness of our scheduler?}

@@

\paragraph{Work stealing} 
When the scheduler finds no work to assign to its workers, it
commences to steal work from other cores using an asynchronous
\texttt{call\_on} active message.  It chooses a victim at random until it
finds one with a non-zero amount of work in its public task queue.  The scheduler steals half of the tasks it finds at the victim.
Work stealing is particularly interesting in Grappa since performance depends
on having many active worker threads on each core. Even if there are many
active threads, if they are all suspended on long-latency operations,
then the core is underutilized. We choose to initiate one outstanding
steal whenever the ready queue is empty there are idle worker threads.

>>>>

The basic unit of execution in \Grappa is a {\em task}. Each task is
represented by a function pointer and its arguments. When a new task is
spawned, it is enqueued until it is ready for execution. Each hardware core
has a single operating system thread pinned to it. A {\em worker} is a
user-level execution thread (stack, \TODO{what else?}) to which tasks are
mapped onto when they are ready to execute.

Once a task is mapped to worker it stays with that worker until it completes. 
During execution, a worker yields control of the processor core whenever it
performs a long-latency operation, allowing the processor to remain busy while
waiting for the operation to complete.
In addition, a programmer can direct
scheduling explicitly via the Grappa API calls shown in
Figure~\ref{fig:scheduling}. To minimize yield overhead, the Grappa scheduler
operates entirely in user-space and does little more than store register state
of one task and load that of another. Context switch times are as low as 38ns
even when switching amongst thousands of tasks.

\Grappa keeps a large number of workers per hardware core, which is at the
core of \Grappa's latency tolerance properties.

\TODO{We should expand on the cleverness of our scheduler?}

% talk about lifetime of a task. activating. blocking. work stealing. look at
% Mark's email. Holt's chat.

@@

\paragraph{Work stealing} 
When the scheduler finds no work to assign to its workers, it commences to
steal tasks from other cores using an asynchronous \texttt{call\_on} active
message. It chooses a victim at random until it finds one with a non-zero
amount of work in its public task queue. The scheduler steals half of the
tasks it finds at the victim. Work stealing is particularly interesting in
Grappa since performance depends on having many active worker threads on each
core. Even if there are many active threads, if they are all suspended on
long-latency operations, then the core is underutilized. We choose to initiate
one outstanding steal whenever the ready queue is empty there are idle worker
threads.
